FOODiE, Fortran Object oriented Ordinary Differential Equations integration library based on Abstract Calculus Pattern

### 2015-10-07 papers/announcement/manuscript.tex

\begin{lstlisting}[firstnumber=1,style=code,caption={integrand type definition},label={list:type_integrand}]
type, abstract :: integrand
  !< Abstract type for building FOODiE ODE integrators.
  contains
    ! public deferred procedures that concrete integrand-field must implement
    procedure(time_derivative),       pass(self), deferred, public:: t
    procedure(update_previous_steps), pass(self), deferred, public:: update_previous_steps
    procedure(previous_step),         pass(self), deferred, public:: previous_step
    ! operators
    procedure(symmetric_operator),   pass(lhs), deferred, public:: integrand_multiply_integrand
    procedure(integrand_op_real),    pass(lhs), deferred, public:: integrand_multiply_real
    procedure(real_op_integrand),    pass(rhs), deferred, public:: real_multiply_integrand
    procedure(symmetric_operator),   pass(lhs), deferred, public:: add
    procedure(symmetric_operator),   pass(lhs), deferred, public:: sub
    procedure(assignment_integrand), pass(lhs), deferred, public:: assign_integrand
    ! operators overloading
    generic, public:: operator(+) => add
    generic, public:: operator(-) => sub
    generic, public:: operator(*) => integrand_multiply_integrand, &
                                     real_multiply_integrand, &
                                     integrand_multiply_real
    generic, public:: assignment(=) => assign_integrand
endtype integrand
\end{lstlisting}

@@

As shown in listing \ref{list:type_integrand}, the number of the deferred type bound procedures that clients must implement into their own concrete extension of the \emph{integrand} ADT is very limited: essentially, there are 3 ODE-specific procedures plus some operators definition constituted by symmetric operators between 2 integrand objects, asymmetric operators between integrand and real numbers (and viceversa) and an assignment statement for the creation of new integrand objects. These procedures are analyzed in the following paragraphs.

@@

\begin{lstlisting}[firstnumber=1,style=code,caption={time derivative procedure interface},label={list:time_derivative}]
function time_derivative(self, n, t) result(dState_dt)
import :: integrand, R_P, I_P
class(integrand),       intent(IN) :: self      !< Integrand field.
integer(I_P), optional, intent(IN) :: n         !< Time level.
real(R_P),    optional, intent(IN) :: t         !< Time.
class(integrand), allocatable      :: dState_dt !< Result of the time derivative function of integrand field.
endfunction time_derivative
\end{lstlisting}

This procedure-function takes three arguments, the first passed as a \emph{type bounded} argument, while the latter are optionals, and it returns an integrand object. The passed dummy argument, \emph{self}, is a polymorphic argument that could be any extensions of the \emph{integrand} ADT. The optional argument \emph{n} indicates the \emph{time level} at which the residuals function must be evaluated: this argument is necessary for multi-step (often referred as multi-level) solvers such as the Adams-Bashforth or leapfrog ones, but it can be omitted for other classes of solvers such the one-step multi-stages family. The optional argument \emph{t} is the \emph{time} at which the residuals function must be computed: it can be omitted in the case the residuals function does not depend directly on time.

@@

\subsubsection{Update previous time steps procedure}

The abstract interface of the \emph{update previous time steps} procedure is shown in listing \ref{list:update_previous_time_steps}.

\begin{lstlisting}[firstnumber=1,style=code,caption={update previous time steps procedure interface},label={list:update_previous_time_steps}]
subroutine update_previous_steps(self, filter, weights)
import :: integrand, R_P
class(integrand),           intent(INOUT) :: self       !< Integrand field.
class(integrand), optional, intent(IN)    :: filter     !< Filter field displacement.
real(R_P),        optional, intent(IN)    :: weights(:) !< Weights for filtering the steps.
endsubroutine update_previous_steps
\end{lstlisting}

This procedure is a necessary complement for using multi-step solvers. As its name suggests, it updates the previous time steps solutions (that must be stored on the care of the clients code). It takes three arguments, the first passed as a \emph{type bounded} argument, while the latter are optionals, and it returns nothing it being a subroutine directly modifying the passed argument. The passed dummy argument, \emph{self}, is a polymorphic argument that could be any extensions of the \emph{integrand} ADT. The optional argument \emph{filter} indicates a \emph{displacement} for filtering the field while the last optional argument \emph{weights} specifies the weights that must applied to the displacement for each time steps filtered.

Let us consider the multi-step scheme expressed by the equation \ref{eq:AB}. Once the solution at time $t^{n+N_s}$ has been computed the previous $N_s-1$ steps must be updated, e.g. an unfiltered update can read as:
\begin{equation}
U\left(t^{n+s}\right) = U\left(t^{n+s+1}\right)\quad s=1,..,N_s-1
\label{eq:previous-steps-update}
\end{equation}
and a filtered update can read as:
\begin{equation}
U\left(t^{n+s}\right) = \Delta * U\left(t^{n+s+1}\right) \omega_s\quad s=1,..,N_s-1
\label{eq:previous-steps-update-filtered}
\end{equation}
where $\Delta$ is the filter displacement and $\omega_s$ the weight of the step $s-th$. This interface implies that the previous time steps must be embedded into the client \emph{integrand} extension. This procedure is automatically called by each multi-step solver after the new step $t^{n+N_s}$ has been computed, thus alleviating client codes from its explicit call.

In the case the client codes will use only one-step solvers this procedure can be defined as simple \emph{empty} subroutine as shown in listing \ref{list:empty_update_previous_time_steps}.

\begin{lstlisting}[firstnumber=1,style=code,caption={empty implementation of the update previous time steps procedure},label={list:empty_update_previous_time_steps}]
subroutine update_previous_steps(self, filter, weights)
class(my_integrand),        intent(INOUT) :: self       !< Integrand field.
class(integrand), optional, intent(IN)    :: filter     !< Filter field displacement.
real(R_P),        optional, intent(IN)    :: weights(:) !< Weights for filtering the steps.
! do nothing
return
endsubroutine update_previous_steps
\end{lstlisting}
where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT.

\subsubsection{Previous step procedure}

The abstract interface of the \emph{previous step} procedure is shown in listing \ref{list:previous_step}.

\begin{lstlisting}[firstnumber=1,style=code,caption={previous step procedure interface},label={list:previous_step}]
function previous_step(self, n) result(previous)
import :: integrand, I_P
class(integrand), intent(IN)  :: self     !< Integrand field.
integer(I_P),     intent(IN)  :: n        !< Time level.
class(integrand), allocatable :: previous !< Selected previous time integrand field.
endfunction previous_step
\end{lstlisting}

This procedure is a necessary complement for using multi-step solvers. It returns the integrand solution at the specified previous time step (level). It takes two arguments and it returns an integrand object. The first argument passed as a \emph{type bounded} argument is the current instance of the integrand field object. The second argument, named \emph{n}, indicates the \emph{time level} at which the solution must be returned.

This procedure is used in multi-step methods like the leapfrog where there is an explicit dependency on also the solution at previous time steps.

For the case in which multi-step solvers are not used, this procedure can have a concrete implementation that does nothing as shown in listing \ref{list:empty_previous_step}.

\begin{lstlisting}[firstnumber=1,style=code,caption={empty implementation of the previous step procedure},label={list:empty_previous_step}]
function previous_step(self, n) result(previous)
class(my_integrand), intent(IN)  :: self     !< Integrand field.
integer(I_P),        intent(IN)  :: n        !< Time level.
class(integrand), allocatable    :: previous !< Selected previous time integrand field.
! do nothing
return
endfunction previous_step
\end{lstlisting}
where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT. Note that, in this case, the returned result is not allocated.

@@

\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of an Adams-Bashforth integrator},label={list:use_ab_solver}]
real               :: times(1:4)
type(my_integrand) :: my_field
call integrator%integrate(U=my_field, Dt=Dt, t=times)
\end{lstlisting}
where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT and \emph{times} are the time at each 4 steps considered for the current one-step-over integration.

@@

\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Adams-Bashforth class},label={list:ab_integrate}]
subroutine integrate(self, U, Dt, t)
class(adams_bashforth_integrator), intent(IN)    :: self ! Actual AB integrator.
class(integrand),                  intent(INOUT) :: U    ! Field to be integrated.
real(R_P),                         intent(IN)    :: Dt   ! Time steps.
real(R_P),                         intent(IN)    :: t(:) ! Times.
integer(I_P)                                     :: s    ! Steps counter.
do s=1, self%steps
  U = U + U%t(n=s, t=t(s)) * (Dt * self%b(s))
enddo
call U%update_previous_steps
return
endsubroutine integrate
\end{lstlisting}

@@

\begin{table}[!ht]
  \centering
  \caption{Oscillation test errors analysis}\label{tab:oscillation}
  \resizebox{0.75\textwidth}{!}{%
  \begin{tabular}{cccccc}
    \toprule
    {\sc Solver} & {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
    \cmidrule{2-6}
    \multirow{6}{*}{Adams-Bashforth 2 steps}
                 &   5000.0        &  0.109E+04    &  0.112E+04    & /             & /             \\
                 &   2500.0        &  0.246E+02    &  0.243E+02    & 5.47          & 5.53          \\
                 &   1250.0        &  0.783E+01    &  0.789E+01    & 1.65          & 1.62          \\
                 &    625.0        &  0.268E+01    &  0.272E+01    & 1.55          & 1.54          \\
                 &    320.0        &  0.976E+00    &  0.990E+00    & 1.51          & 1.51          \\
                 &    100.0        &  0.171E+00    &  0.173E+00    & 1.50          & 1.50          \\
    \cmidrule{2-6}
    \multirow{6}{*}{Adams-Bashforth 3 steps}
                 &   5000.0        &  0.816E+01    &  0.809E+01    & /             & /             \\
                 &   2500.0        &  0.266E+01    &  0.264E+01    & 1.62          & 1.62          \\
                 &   1250.0        &  0.120E+01    &  0.120E+01    & 1.15          & 1.14          \\
                 &    625.0        &  0.545E+00    &  0.543E+00    & 1.14          & 1.14          \\
                 &    320.0        &  0.223E+00    &  0.222E+00    & 1.34          & 1.34          \\
                 &    100.0        &  0.418E-01    &  0.416E-01    & 1.44          & 1.44          \\
    \cmidrule{2-6}
    \multirow{6}{*}{Adams-Bashforth 4 steps}
                 &   5000.0        &  0.384E+09    &  0.380E+09    & /             & /             \\
                 &   2500.0        &  0.871E+01    &  0.869E+01    & 25.39         & 25.38         \\
                 &   1250.0        &  0.314E+01    &  0.312E+01    &  1.47         &  1.48         \\
                 &    625.0        &  0.111E+01    &  0.111E+01    &  1.49         &  1.49         \\
                 &    320.0        &  0.409E+00    &  0.407E+00    &  1.50         &  1.50         \\
                 &    100.0        &  0.714E-01    &  0.711E-01    &  1.50         &  1.50         \\

    \cmidrule{2-6}
    \multirow{6}{*}{Leapfrog RAW-filtered}
                 &   5000.0        &  0.151E+02    &  0.152E+02    & /             & /             \\
                 &   2500.0        &  0.500E+01    &  0.504E+01    & 1.60          & 1.59          \\
                 &   1250.0        &  0.190E+01    &  0.193E+01    & 1.40          & 1.39          \\
                 &    625.0        &  0.181E+01    &  0.181E+01    & 0.07          & 0.09          \\
                 &    320.0        &  0.157E+01    &  0.157E+01    & 0.21          & 0.21          \\
                 &    100.0        &  0.100E+01    &  0.100E+01    & 0.39          & 0.39          \\
    \cmidrule{2-6}
    \multirow{6}{*}{low storage Runge-Kutta 5 stages}
                 &   5000.0        &  0.120E+00    &  0.122E+00    & /             & /             \\
                 &   2500.0        &  0.106E-01    &  0.107E-01    & 3.51          & 3.51          \\
                 &   1250.0        &  0.935E-03    &  0.947E-03    & 3.50          & 3.50          \\
                 &    625.0        &  0.826E-04    &  0.836E-04    & 3.50          & 3.50          \\
                 &    320.0        &  0.793E-05    &  0.803E-05    & 3.50          & 3.50          \\
                 &    100.0        &  0.135E-06    &  0.137E-06    & 3.50          & 3.50          \\
    \cmidrule{2-6}
    \multirow{6}{*}{TVD/SSP Runge-Kutta 2 stages}
                 &   5000.0        &  0.316E+02    &  0.319E+02    & /             & /             \\
                 &   2500.0        &  0.892E+01    &  0.894E+01    & 1.83          & 1.84          \\
                 &   1250.0        &  0.301E+01    &  0.305E+01    & 1.57          & 1.55          \\
                 &    625.0        &  0.106E+01    &  0.107E+01    & 1.51          & 1.51          \\
                 &    320.0        &  0.387E+00    &  0.392E+00    & 1.50          & 1.50          \\
                 &    100.0        &  0.676E-01    &  0.685E-01    & 1.50          & 1.50          \\
    \cmidrule{2-6}
    \multirow{6}{*}{TVD/SSP Runge-Kutta 3 stages}
                 &   5000.0        &  0.255E+01    &  0.252E+01    & /             & /             \\
                 &   2500.0        &  0.523E+00    &  0.516E+00    & 2.28          & 2.29          \\
                 &   1250.0        &  0.944E-01    &  0.931E-01    & 2.47          & 2.47          \\
                 &    625.0        &  0.167E-01    &  0.165E-01    & 2.50          & 2.50          \\
                 &    320.0        &  0.314E-02    &  0.310E-02    & 2.50          & 2.50          \\
                 &    100.0        &  0.171E-03    &  0.169E-03    & 2.50          & 2.50          \\
    \cmidrule{2-6}
    \multirow{6}{*}{TVD/SSP Runge-Kutta 5 stages}
                 &   5000.0        &  0.139E+00    &  0.141E+00    & /             & /             \\
                 &   2500.0        &  0.122E-01    &  0.124E-01    & 3.50          & 3.50          \\
                 &   1250.0        &  0.108E-02    &  0.110E-02    & 3.50          & 3.50          \\
                 &    625.0        &  0.956E-04    &  0.969E-04    & 3.50          & 3.50          \\
                 &    320.0        &  0.937E-05    &  0.949E-05    & 3.47          & 3.47          \\
                 &    100.0        &  0.512E-06    &  0.519E-06    & 2.50          & 2.50          \\
    \bottomrule
  \end{tabular}}
\end{table}

@@

{\color{red}This class of solvers does not behaves as expected.

The expected observed orders of accuracy should tend to 2, 3 and 4 for the corresponding Adams-Bashforth 2, 3 and 4 steps, but the observed accuracy tend to 1.5 for all solvers. Is there a bug? Is the Adams-Bashforth not suited for the Oscillation test?

Figure \ref{fig:results-oscillation-adams-bashforth} emphasizes instability generation for some pairs scheme-$\Delta t$: odd-even phase/amplitude phenomena are present. A deep insight from whom is familiar with multi-step method is welcome.

Note that in the summary.log file into error analysis directory I have computed also the product of $f*\Delta t$ in order to verify when the instability arises.

For AB-2 and AB-4 the solution with $\Delta t=5000 \rightarrow f*\Delta t=0.5$ results to be unstable (the computational modes are not dumped) while for AB-3 it is stable, but the (physical mode) amplitude is dumped resulting into large errors. This is consistent with Duran results, where AB-3 is said to be unstable for $ f*\Delta t>0.724$.

The AB-2 and AB-4 solutions show a phase error that decreases as the time resolution increases. The AB-3 has no phase error and the computational modes are dumped for all $\Delta t$ exercised.
}

>>>>

\begin{lstlisting}[firstnumber=1,style=code,caption={integrand type definition},label={list:type_integrand}]
type, abstract :: integrand
  !< Abstract type for building FOODiE ODE integrators.
  contains
    ! public deferred procedures that concrete integrand-field must implement
    procedure(time_derivative),       pass(self), deferred, public:: t
    ! operators
    procedure(symmetric_operator),   pass(lhs), deferred, public:: integrand_multiply_integrand
    procedure(integrand_op_real),    pass(lhs), deferred, public:: integrand_multiply_real
    procedure(real_op_integrand),    pass(rhs), deferred, public:: real_multiply_integrand
    procedure(symmetric_operator),   pass(lhs), deferred, public:: add
    procedure(symmetric_operator),   pass(lhs), deferred, public:: sub
    procedure(assignment_integrand), pass(lhs), deferred, public:: assign_integrand
    ! operators overloading
    generic, public:: operator(+) => add
    generic, public:: operator(-) => sub
    generic, public:: operator(*) => integrand_multiply_integrand, &
                                     real_multiply_integrand, &
                                     integrand_multiply_real
    generic, public:: assignment(=) => assign_integrand
endtype integrand
\end{lstlisting}

@@

As shown in listing \ref{list:type_integrand}, the number of the deferred type bound procedures that clients must implement into their own concrete extension of the \emph{integrand} ADT is very limited: essentially, there are 1 ODE-specific procedure plus some operators definition constituted by symmetric operators between 2 integrand objects, asymmetric operators between integrand and real numbers (and viceversa) and an assignment statement for the creation of new integrand objects. These procedures are analyzed in the following paragraphs.

@@

\begin{lstlisting}[firstnumber=1,style=code,caption={time derivative procedure interface},label={list:time_derivative}]
function time_derivative(self, t) result(dState_dt)
import :: integrand, R_P, I_P
class(integrand),       intent(IN) :: self      !< Integrand field.
real(R_P),    optional, intent(IN) :: t         !< Time.
class(integrand), allocatable      :: dState_dt !< Result of the time derivative function of integrand field.
endfunction time_derivative
\end{lstlisting}

This procedure-function takes two arguments, the first passed as a \emph{type bounded} argument, while the latter is optional, and it returns an integrand object. The passed dummy argument, \emph{self}, is a polymorphic argument that could be any extensions of the \emph{integrand} ADT. The optional argument \emph{t} is the \emph{time} at which the residuals function must be computed: it can be omitted in the case the residuals function does not depend directly on time.

@@

@@

\begin{lstlisting}[firstnumber=1,style=code,caption={example of usage of an Adams-Bashforth integrator},label={list:use_ab_solver}]
real               :: times(1:4)
type(my_integrand) :: my_field
type(my_integrand) :: previous(1:4)
call integrator%integrate(U=my_field, previous=previous, Dt=Dt, t=times)
\end{lstlisting}
where \emph{my\_integrand} is a concrete (valid) extension of \emph{integrand} ADT, \emph{times} are the time at each 4 steps considered for the current one-step-over integration and \emph{previous} are the memory registers where previous time steps solutions are saved.

@@

{\color{red} To be modified with changed API}

\begin{lstlisting}[firstnumber=1,style=code,caption={implementation of the \emph{integrate} method of explicit Adams-Bashforth class},label={list:ab_integrate}]
subroutine integrate(self, U, previous, Dt, t)
class(adams_bashforth_integrator), intent(IN)    :: self ! Actual AB integrator.
class(integrand),                  intent(INOUT) :: U    ! Field to be integrated.
real(R_P),                         intent(IN)    :: Dt   ! Time steps.
real(R_P),                         intent(IN)    :: t(:) ! Times.
integer(I_P)                                     :: s    ! Steps counter.
do s=1, self%steps
  U = U + U%t(n=s, t=t(s)) * (Dt * self%b(s))
enddo
call U%update_previous_steps
return
endsubroutine integrate
\end{lstlisting}

@@

\begin{table}[!ht]
  \centering
  \caption{Oscillation test errors analysis}\label{tab:oscillation}
  \resizebox{0.75\textwidth}{!}{%
  \begin{tabular}{cccccc}
    \toprule
    {\sc Solver} & {\sc Time Step} & {\sc Error X} & {\sc Error Y} & {\sc Order X} & {\sc Order Y} \\
    \cmidrule{2-6}
    \multirow{6}{*}{Adams-Bashforth 2 steps}
                 &   5000.0        &  0.596E+03    &  0.583E+03    & /             & /             \\
                 &   2500.0        &  0.221E+02    &  0.218E+02    & 4.75          & 4.74          \\
                 &   1250.0        &  0.764E+01    &  0.769E+01    & 1.53          & 1.50          \\
                 &    625.0        &  0.265E+01    &  0.268E+01    & 1.53          & 1.52          \\
                 &    320.0        &  0.968E+00    &  0.981E+00    & 1.51          & 1.50          \\
                 &    100.0        &  0.169E+00    &  0.171E+00    & 1.50          & 1.50          \\
    \cmidrule{2-6}
    \multirow{6}{*}{Adams-Bashforth 3 steps}
                 &   5000.0        &  0.857E+01    &  0.854E+01    & /             & /             \\
                 &   2500.0        &  0.391E+01    &  0.386E+01    & 1.13          & 1.14          \\
                 &   1250.0        &  0.825E+00    &  0.814E+00    & 2.24          & 2.25          \\
                 &    625.0        &  0.150E+00    &  0.148E+00    & 2.46          & 2.46          \\
                 &    320.0        &  0.282E-01    &  0.278E-01    & 2.49          & 2.49          \\
                 &    100.0        &  0.154E-02    &  0.152E-02    & 2.50          & 2.50          \\
    \cmidrule{2-6}
    \multirow{6}{*}{Adams-Bashforth 4 steps}
                 &   5000.0        &  0.128E+07    &  0.143E+07    & /             & /             \\
                 &   2500.0        &  0.106E+01    &  0.107E+01    & 20.21         & 20.34         \\
                 &   1250.0        &  0.967E-01    &  0.981E-01    &  3.45         &  3.45         \\
                 &    625.0        &  0.859E-02    &  0.871E-02    &  3.49         &  3.49         \\
                 &    320.0        &  0.827E-03    &  0.838E-03    &  3.50         &  3.50         \\
                 &    100.0        &  0.141E-04    &  0.143E-04    &  3.50         &  3.50         \\

    \cmidrule{2-6}
    \multirow{6}{*}{Leapfrog RAW-filtered}
                 &   5000.0        &  0.156E+02    &  0.156E+02    & /             & /             \\
                 &   2500.0        &  0.855E+01    &  0.852E+01    & 0.86          & 0.87          \\
                 &   1250.0        &  0.303E+01    &  0.305E+01    & 1.50          & 1.48          \\
                 &    625.0        &  0.107E+01    &  0.108E+01    & 1.51          & 1.50          \\
                 &    320.0        &  0.390E+00    &  0.395E+00    & 1.50          & 1.50          \\
                 &    100.0        &  0.685E-01    &  0.692E-01    & 1.50          & 1.50          \\
    \cmidrule{2-6}
    \multirow{6}{*}{low storage Runge-Kutta 5 stages}
                 &   5000.0        &  0.120E+00    &  0.122E+00    & /             & /             \\
                 &   2500.0        &  0.106E-01    &  0.107E-01    & 3.51          & 3.51          \\
                 &   1250.0        &  0.935E-03    &  0.947E-03    & 3.50          & 3.50          \\
                 &    625.0        &  0.826E-04    &  0.836E-04    & 3.50          & 3.50          \\
                 &    320.0        &  0.793E-05    &  0.803E-05    & 3.50          & 3.50          \\
                 &    100.0        &  0.135E-06    &  0.137E-06    & 3.50          & 3.50          \\
    \cmidrule{2-6}
    \multirow{6}{*}{TVD/SSP Runge-Kutta 2 stages}
                 &   5000.0        &  0.316E+02    &  0.319E+02    & /             & /             \\
                 &   2500.0        &  0.892E+01    &  0.894E+01    & 1.83          & 1.84          \\
                 &   1250.0        &  0.301E+01    &  0.305E+01    & 1.57          & 1.55          \\
                 &    625.0        &  0.106E+01    &  0.107E+01    & 1.51          & 1.51          \\
                 &    320.0        &  0.387E+00    &  0.392E+00    & 1.50          & 1.50          \\
                 &    100.0        &  0.676E-01    &  0.685E-01    & 1.50          & 1.50          \\
    \cmidrule{2-6}
    \multirow{6}{*}{TVD/SSP Runge-Kutta 3 stages}
                 &   5000.0        &  0.255E+01    &  0.252E+01    & /             & /             \\
                 &   2500.0        &  0.523E+00    &  0.516E+00    & 2.28          & 2.29          \\
                 &   1250.0        &  0.944E-01    &  0.931E-01    & 2.47          & 2.47          \\
                 &    625.0        &  0.167E-01    &  0.165E-01    & 2.50          & 2.50          \\
                 &    320.0        &  0.314E-02    &  0.310E-02    & 2.50          & 2.50          \\
                 &    100.0        &  0.171E-03    &  0.169E-03    & 2.50          & 2.50          \\
    \cmidrule{2-6}
    \multirow{6}{*}{TVD/SSP Runge-Kutta 5 stages}
                 &   5000.0        &  0.139E+00    &  0.141E+00    & /             & /             \\
                 &   2500.0        &  0.122E-01    &  0.124E-01    & 3.50          & 3.50          \\
                 &   1250.0        &  0.108E-02    &  0.110E-02    & 3.50          & 3.50          \\
                 &    625.0        &  0.956E-04    &  0.969E-04    & 3.50          & 3.50          \\
                 &    320.0        &  0.937E-05    &  0.949E-05    & 3.47          & 3.47          \\
                 &    100.0        &  0.512E-06    &  0.519E-06    & 2.50          & 2.50          \\
    \bottomrule
  \end{tabular}}
\end{table}

@@

The expected observed orders of accuracy for the Adams-Bashforth solvers using 2, 3 and 4 time steps tend to 1.5, 2.5 and 3.5 that are in agreement with the expected formal order of 2, 3 and 4, respectively.

Figure \ref{fig:results-oscillation-adams-bashforth} emphasizes the instability generation for some pairs steps number/$\Delta t$. The 2 and 4 steps solutions are instable for $\Delta t=5000 \rightarrow f*\Delta t=0.5$. On the contrary, the 3 steps solution is stable, but the amplitude is dumped and the solution vanishes as the integration proceeds. The 2 and 4 steps solutions show a phase error that decreases as the time resolution increases, whereas 3 steps solution has null phase error.
